# Reproducible Research Assignment 1
### Author: Reem Sarrif

This R Markdown document contains the code chunks and explanatory text required for the Reproducible Research Assignment 1.

##Loading and Preprocessing the data
In this part we start by unzipping the file to extract the "activity.csv" file, 
then we read this into the data frame "activity".

```{r}
  unzip("activity.zip")
  activity <- read.csv("activity.csv")
  str(activity)
```

##What is the mean total number of steps taken per day?
The total number of steps taken per day is calculated, and a histogram is displayed to show the total number of steps taken each day.

```{r}
  ActivitySum <- aggregate(steps ~ date, data=activity, FUN="sum", na.rm=TRUE)

  hist(ActivitySum$steps, col="red")
```

Next week calculate the mean and median for total number of steps taken per day.
Note: the code chunk for this part is hidden and results shown below.

```{r activitysummaries, echo=FALSE, results="hide"}
  dailymean <- mean(ActivitySum$steps)
  dailymean <- format(dailymean, digits=7, nsmall=2)
  dailymedian <- median(ActivitySum$steps)
  dailymedian <- format(dailymedian, digits=7, nsmall=2)
```

For the total number of steps taken per day, the mean is `r dailymean` and the median is `r dailymedian`


##What is the average daily activity pattern?
Here we calculate the mean for number of steps taken per 5-minute interval across all days.  Results plotted as time series plot.

```{r}
  ActivitySum <- aggregate(steps ~ interval, data=activity, FUN="mean", na.rm=TRUE)
  with(ActivitySum, plot(interval,steps, type="l"))
```

Next we find the interval with the maximum average number of steps.
Note: the code for this chunk is hidden and findings displayed below.

```{r maxinterval, echo=FALSE, results="hide"}
  maxinterval <- ActivitySum$interval[ActivitySum$steps==max(ActivitySum$steps)]
```

The 5-minute interval on average across all days that contains the maximum number of steps is interval `r maxinterval`


##Imputing missing values
```{r nacount, echo=FALSE, results="hide"}
  nacount <- nrow(activity[activity$steps=="NA",])
```

The total number of missing values in the dataset is `r nacount`

```{r}
  ActivitySum <- aggregate(steps ~ date, data=activity, FUN="mean", na.rm=TRUE)
  activityImp <- merge(activity, ActivitySum, by="date")
  activityImp$steps <- ifelse(is.na(activityImp$steps.x)==TRUE,activityImp$steps.y,activityImp$steps.x)
  head(activity)
  head(activityImp)
  hist(activityImp$steps, col="red")
  ActivitySumImp <- aggregate(steps ~ date, data=activityImp, FUN="sum", na.rm=TRUE)
  
```


```{r activitysummaries1, echo=FALSE, results="hide"}
  dailymeanImp <- mean(ActivitySumImp$steps)
  dailymeanImp <- format(dailymeanImp, digits=7, nsmall=2)
  dailymedianImp <- median(ActivitySumImp$steps)
  dailymedianImp <- format(dailymedianImp, digits=7, nsmall=2)
```

For the total number of steps taken per day, the mean is `r dailymeanImp` and the median is `r dailymedianImp`

TODO

##Are there differences in activity patterns between weekdays and weekends?

```{r}
  activity$daytype <- ifelse(weekdays(as.Date(activity$date)) %in% c("Saturday","Sunday"),"weekend","weekday")
  library(lattice)
  xyplot(steps ~ interval | daytype, data=activity, layout = c(2,1), type="l")
```



